Fitting metabolic models to dissolved oxygen data: the Estuarine BAyesian Single-station Estimation method

### 2023-05-03 manu-draft.qmd
update fig and text for priorsumcomp

@fig-priorsumcomp provides an alternative summary of how the prior distributions influenced the ability of EBASE to reproduce the synthetic parameters.  Median NSE values across all models and each optimization period for the different subsets of the prior combinations are shown.  For example, the bottom left panel with the x-axis value for the $a$ parameter shows the median NSE values for all models using a one-day optimization period where the $\sigma$ value for the $a$ prior distribution was low or high.  The size of the points indicate the inter-quartile range of NSE values for each subset.  As such, each set of points denoted by the x-axis parameters in each panel include median NSE values for all models by optimization period, but subset by the low/high $\mu$ or $\sigma$ values for each parameter. Although most of the median NSE values are less than zero, indicating a poor ability of EBASE to reproduce the parameters, patterns emerge as to which characteristics of the prior lead to robust estimates.  For example, large differences in the median NSE values are observed for low and high $\sigma$ for the $a$ and $r$ parameters, with positive median NSE values and smaller IQR values when using a high $\sigma$ value using the 7 and 30 day optimization periods.  A similar, but smaller effect is seen for the $\sigma$ value of the prior for the $b$ parameter.  A conclusion similar to that for @fig-priorcomp is that large standard deviations in the prior distributions lead to more accurate estimates of metabolic parameters.  Conversely, model results are less sensitive to low or high values for $\mu$ as the spread between median NSE values is less than those for $\sigma$.  This suggests that EBASE model sensitivity is most influenced by $\sigma$ of the prior distributions, i.e., larger values produce more accurate and less variable estimates regardless of $\mu$.

@@

![Summaries of median Nash-Sutcliffe Efficiency (NSE) values across all parameters for the comparisons of EBASE and synthetic model outputs separated by different combinations of prior values. Each result on the x-axis represents the median values for which the combination prior applies, e.g., all results from @fig-priorcomp where $\mu$ is high or low for the $a$ parameter using a one day optimization period.  Point size represents the inter-quartile range of median NSE values for each result.](figs/priorsumcomp.png){#fig-priorsumcomp}

>>>>

@fig-priorsumcomp provides an alternative summary of how the prior distributions influenced the ability of EBASE to reproduce the synthetic parameters.  Median NSE values across all models, parameters, and each optimization period for the different subsets of the prior combinations are shown relative to the grand median NSE value across all comparisons.  For example, the bottom left panel with the x-axis value for the $a$ parameter shows the difference in the median NSE values for all models using a one-day optimization period where the $\sigma$ value for the $a$ prior distribution was low or high.  The size of the points indicate the inter-quartile range of NSE values for each subset.  As such, each set of points denoted by the x-axis parameters in each panel include median NSE values for all models by optimization period, but subset by the low/high $\mu$ or $\sigma$ values for each parameter. Positive values indicate relative performance greater than the overall median NSE values and negative values indicate relative performance worse than the overall median.  Dissolved oxygen was generally modeled with high accuracy for all priors and it was excluded to not skew the assessment.  The patterns suggest that EBASE results are more sensitive to changes in the $\sigma$ parameter as shown by the larger spread in the median results compared to those for $\mu$. This result is especially true for the priors of the $a$ and $r$ parameters.  Conversely, model results are less sensitive to low or high values for $\mu$ as the spread between median NSE values is less than those for $\sigma$.  This suggests that EBASE model sensitivity is most influenced by $\sigma$ of the prior distributions, i.e., larger values produce more accurate and less variable estimates regardless of $\mu$.  Patterns by optimization period were not apparent.

@@

![Summaries of median Nash-Sutcliffe Efficiency (NSE) values across all parameters for the comparisons of EBASE and synthetic model outputs separated by different combinations of prior values. Summaries are shown relative to the difference from the grand median across all comparisons.  Each result on the x-axis represents the median values for which the combination prior applies, e.g., all results from @fig-priorcomp where $\mu$ is high or low for the $a$ parameter using a one day optimization period.  Point size represents the inter-quartile range of median NSE values for each result.](figs/priorsumcomp.png){#fig-priorsumcomp}
